---
title: "SADBank - Standard Area Diagrams of Plant Disease Severity"
output: 
   flexdashboard::flex_dashboard:
   vertical_layout: scroll
bibliography: sad.bib   
   
---

<script src="https://use.fontawesome.com/682f97eaed.js"></script>

<style type="text/css">
body{
   font-size: 16px;
      }
      
      </style>

Summary tables {data-icon="fa-table"}
=====================================


```{r, include=FALSE}
# reading the data from file
library(ggplot2)
library(plotly)
library(readr)
library(plyr)
library(dplyr)
dat_sad <- read_csv ("dat_sad.csv")
attach(dat_sad)
dat_sad$citation <- paste(author, pub_year)
dat_sad$study_diag <- with(dat_sad, interaction(study, diag_id))
dat_sad$citation_diag <- with(dat_sad, interaction(citation, diag_id))
detach(dat_sad)
# aggregate by individidual studies
dat_sad <- group_by(dat_sad, study)
dat_sad1 <- filter(dat_sad, row_number()==1 )
```






Column {.tabset }
-------------------------------------
 


### Study information

```{r, include=FALSE}

dat_sad2 <- ungroup(dat_sad1)
table1 <- select(dat_sad2, pub_year, crop, pathogen, pathogen_group, journal, country, lang, pub_type, citation, link)
table1 <- arrange(table1, -pub_year)
```

```{r}
library(magrittr)
library(DT)

datatable(table1,
          escape = FALSE,
          class = 'cell-border stripe',
          rownames = FALSE, colnames = c('Publication year','Crop', 'Pathogen', 'Pathogen group', 'Journal', 'Country', 'Language', 'Publication type',  'Reference', 'Link'), 
          options = list(
  pageLength = 25,
  fontSize = 12,
  lengthMenu = c(10, 25, 50, 100 )
))%>%
formatStyle(
  c('crop', 'pathogen', 'pathogen_group', 'pub_year','journal', 'country', 'lang', 'pub_type', 'citation', 'link'),
  fontSize = '90%')

```

   
### SAD characteristics

```{r, include=FALSE}
library(plyr)
library(dplyr)
dat_sad2 <- ungroup(dat_sad1)
table1 <- select(dat_sad2, pub_year, crop, dis_name, pathogen, n_diag, organ, diag_type, col_type, n_diagrams, min_sev, max_sev, citation, link)

table1 <- arrange(table1, -pub_year)
```

```{r}


datatable(table1, 
          escape = FALSE,
          class = 'cell-border stripe',
        rownames = FALSE, colnames = c('Publication year','Crop', 'Disease name', 'Pathogen', 'N. of sets','Organ', 'Design type', 'Diagram color', 'N. of diagrams', 'Min. Severity', 'Max. Severity', 'Reference', "Link"), 
          options = list(
  pageLength = 25,
  fontSize = 12,
  lengthMenu = c(10, 25, 50, 100 )
))%>%
formatStyle(
  c('pub_year', 'crop','dis_name','pathogen', 'n_diag', 'organ', 'diag_type', 'col_type', 'n_diagrams', 'min_sev', 'max_sev', 'citation', 'link'),
  fontSize = '90%')

```


### SAD validation approach 

```{r, include=FALSE}
library(plyr)
library(dplyr)
dat_sad2 <- ungroup(dat_sad1)
table1 <- select(dat_sad2, pub_year, crop, dis_name, pathogen, validation, stat, coefs, n_valid, citation)

table1 <- arrange(table1, -pub_year)
```

```{r}

datatable(table1, 
          class = 'cell-border stripe',
                  rownames = FALSE, colnames = c('Publication year','Crop', 'Disease name', 'Pathogen', 'Validation approach', 'Statistics', 'Reported coefficients', 'N. of specimens assessed', 'Reference'), 
          options = list(
  pageLength = 25,
  fontSize = 12,
  lengthMenu = c(10, 25, 50, 100 )
))%>%
formatStyle(
  c('pub_year', 'crop','dis_name','pathogen', 'validation', 'stat', 'coefs', 'n_valid', 'citation'),
  fontSize = '90%')

```

Summary Plots {data-icon="fa-bar-chart"}
=====================================



Column {.tabset}
----------------------------------

### Articles/year


```{r, echo=FALSE}
library(ggplot2)
library(plotly)
g_artigos <- ggplot(dat_sad1, aes(factor(pub_year), fill=lang)) + geom_bar() +
         scale_y_continuous(breaks=seq(0, 12, 1), expand=c(0.02, 0))+
       xlab("Year")+ ylab("Number of articles")+
  theme_minimal()+
       theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none")+
         scale_fill_manual(values = c("steelblue","brown", "limegreen"))+
  labs(fill = "Language")
      
(gg <- ggplotly(g_artigos))
``` 

### N. of samples

```{r}
n_dev <- subset(dat_sad2, n_dev <1000, select = n_dev)
g_nsamples <- ggplot(n_dev, aes(x=n_dev))+
geom_histogram(binwidth=50, color = "white", fill = "SteelBlue")+ 
 xlab("Number of samples")+
 ylab("Frequency")+
 scale_x_continuous(breaks=c(0, 50, 100, 150, 200, 250, 300, 350, 400, 450, 500))+
  theme_minimal()
(gg <- ggplotly(g_nsamples, tooltip = c("count")))


```


### Software

```{r}
library(RColorBrewer)
dat_sad1$sev_quant2 <- with(dat_sad1, reorder(sev_quant, sev_quant, function(x) length(x)))
g_soft <- ggplot(dat_sad1, aes(sev_quant2, fill=sev_quant2)) +
            geom_bar() + coord_flip()+
            scale_y_continuous(breaks=seq(0, 100, 2),
                               expand=c(0, 0))+
            xlab("")+ ylab("Number of studies")+
  theme_minimal()+
            theme(legend.position="none")+
 scale_fill_hue("clarity", direction = -1)
           
(gg <- ggplotly(g_soft, tooltip = c("count")))

```

### N. of diagrams
```{r}


g_ndiag<- ggplot(dat_sad1, aes(x=n_diagrams))+
geom_histogram( binwidth=1, color = "white", fill = "steelblue")+ 
 xlab("Number of diagrams in the SAD")+
 ylab("Frequency")+
 scale_x_continuous(breaks=c(1:12))+
theme_minimal()
(gg <- ggplotly(g_ndiag, tooltip = c("count")))
```

### Min. sev

```{r}
g_minsev <- ggplot(dat_sad2, aes(x=min_sev))+
geom_histogram(fill = "steelblue", color = "white")+ 
 xlab("Minimum severity (%)")+
 ylab("Frequency")+
 scale_x_continuous(breaks=c(0:11))+
  theme_minimal()
(gg <- ggplotly(g_minsev))
```

### Max. sev

```{r}
g_maxsev <- ggplot(dat_sad2, aes(x=max_sev))+
geom_histogram(fill = "steelblue", color = "white")+ 
 xlab("Maximum severity (%)")+
 ylab("Frequency")+
 scale_x_continuous(breaks=c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100 ))+
 theme_minimal()
(gg <- ggplotly(g_maxsev, tooltip = c("count")))
```

### SAD color

```{r}
dat_sad2$col_type2 <- with(dat_sad2, reorder(col_type, col_type, function(x) length(x)))
g_color <- ggplot(dat_sad2, aes(col_type2, fill=diag_type)) +
            geom_bar() + coord_flip()+
            scale_y_continuous(breaks=seq(0, 100, 5), expand=c(0.02, 0))+
            xlab("")+ ylab("Number of SADs")+
 theme_minimal()+
            theme(legend.position = "none")+ scale_fill_manual(values = c("steelblue", "limegreen"))
 (gg <- ggplotly(g_color))
```

### Max sev x ndiag

```{r}
g_rel1 <- ggplot(dat_sad2, aes(max_sev, n_diagrams))+
geom_point(aes(alpha=0.5), size=3, colour = "limegreen",
            position = position_jitter(width=.2))+
 stat_smooth(colour="steelblue", size=0.75, alpha=0.1)+
  theme_minimal()+
 theme(legend.position = "none")+
 scale_y_continuous(breaks=c(1:12 ))+
 ylab("Number of diagrams")+
 xlab("Maximum severity (%)")
(gg <- ggplotly(g_rel1))


```


### Country

```{r}
library(plotly)


df <- read.csv('country.csv')

# light grey boundaries
l <- list(color = toRGB("grey"), width = 0.5)

# specify map projection/options
g <- list(
  showframe = FALSE,
  showcoastlines = TRUE)


plot_ly(df, z = SAD, text = COUNTRY, locations = CODE, type = 'choropleth', showscale = FALSE,
        color = SAD, colors = c("white", "steelblue"), marker = list(line = l), title = '') %>%
  layout(title = '<br>Country of the authorship of studies</br> on the development and validation of SADs',
         geo = g)

```

Column {.tabset data-width=300}
----------------------------------

### Variable descriptions


**Articles/year**: 
Number of articles per year and language of the publication where EN = english, ES = spanish and PT = Portuguese

**N. of samples**:
Frequency of the number of specimens (leaf, fruit, stalk, etc.) collected in the field or experimental plots which were analyzed to determine the range of disease severity.

**Software**:
Frequency of the approach used to determine the actual severity values in the samples such as image analysis tools or instruments such as leaf area meater.

**N. of diagrams**: 
Number of diagrams depicted in an SAD set.

**Min sev**:
Frequency of minimum non-zero severity depicted in the SAD.

**Max sev**: 
Frequency of maximum severity depicted in the SAD.

**SAD color**:
Frequency of the color scheme used to depict the SAD such as black-white, 2-color, true-color, etc.

**Max sev x ndiag**:
Relationship between the number of diagrams  and maximum severity depicted in the SAD.

**Country**: 
Country of the research group that conducted the study.

### Help and credits

**What is SADBank?**

The **SADBank** is a repository of SADs that supplements a systematic review article published in Phytopathology. Currently, the number of SADs added to the bank is

`nrow(table1` 

Del Ponte EM, Pethybridge SJ, Machado FJ, Michereff SJ. Standard area diagrams for aiding disease severity estimation: systematic review and meta-analysis of the gains in precision. Phytopathology.

**What is an SAD?**

Standard area diagrams (SADs) are a set of pictorial representation of disease symptoms/signs in a plant organ. Each diagram of the set depicts a percent diseased area. They are also called diagrammatic scales or disease diagrams. They are designed in a range of styles, but most commonly as drawing or photographs of diseased organs represented in black-white or true-color. The number of diagrams in the set varies according to the disease. 

**How are they used?**

SADs are tools used during or prior to visual assessments of disease severity to "calibrate" a rater's eye with the aim to improve accuracy, precision and reliability of the estimates. During estimation, a rater assigns a percent value that should be as close as much to actual values. The diagrams in an SAD set just serves as a guide for interpolation. In most studies, SADs have proven of value especially for innately inaccurate raters and certain symptoms/signs that are more difficult to assess.





